---
title: "Background on the federal election in Germany (2017)"
author: "Franziska Löw"
date: "`r format(Sys.Date())`"
output: 
  html_document:
    code_folding: hide
---

```{r include=FALSE}
library(rvest)
library(stringr)  
library(stringi)
library(rebus)     
library(lubridate)
library(tidyverse)
library(zoo)

rm(list=ls())
```

## Evaluation of Sunday surveys

For each party I take the survey results of all institutes at the same time. To calculate a smooth line for each party on each day, I take the moving average within 15 days (7 before the day, 7 after the day, and the day itself). 

Since the institutions do not all publish new values on the same days, the overall temporal accuracy is higher than the weekly accuracy. The data source is [wahlrecht.de](https://www.wahlrecht.de/).

```{r}
get_institute_data <- function(institute) {
  
  url <- paste0(baseurl,institute,".htm")
  
  scrape <- read_html(url)
  
  ## Get colnames ##
  scrape %>% 
    html_node("table.wilko") %>%
    html_node("thead") %>%
    html_text() -> colnames
  
  colnames <- strsplit(colnames,'\\n+')[[1]]
  
  ## Get table rows ##
  scrape %>% 
    html_node("table.wilko") %>%
    html_node("tbody") %>%
    html_nodes("tr") %>%
    html_text() -> rows
  
  out = NULL
  for (row in rows) {
    new_row <- strsplit(row,'\\n+')[[1]]
    out <- rbind(out,new_row)
    }
  
  df <- as.tibble(out)
  names(df) <- colnames 
  df$institute <- institute
  return(df)
}
```

```{r eval=FALSE, include=FALSE}
baseurl <- "https://www.wahlrecht.de/umfragen/"
colnames <- c("Date","CDU/CSU","SPD","GRÜNE","FDP","LINKE","AfD","Sonstige","Befragte","Zeitraum")

institutes <- c("allensbach","insa","emnid","forsa","politbarometer","gms","dimap")

df1 <- get_institute_data(institutes[1])

df1 <- df1[c(1,3,4,5,6,7,8,9,13)]

df1 <- df1 %>% 
  gather(key = "party", value = pollvalue, ` CDU/CSU`:` Sonstige`) %>%
  transmute(
    date = as.Date(` `, format = "%d.%m.%Y"),
    party = gsub("\\s","",party),
    pollvalue1 = gsub(" %","",pollvalue),
    pollvalue = as.numeric(gsub(",",".",pollvalue1)),
    institute = institute
  ) %>%
  select(-pollvalue1)

df2 <- get_institute_data(institutes[2])

df2 <- df2[c(1,3,4,5,6,7,8,9,10,11,15)]

df2 <- df2 %>% 
  gather(key = "party", value = pollvalue, ` CDU/CSU`:` Sonstige`) %>%
  transmute(
    date = as.Date(Datum, format = "%d.%m.%Y"),
    party = gsub("\\s","",party),
    pollvalue1 = gsub(" %","",pollvalue),
    pollvalue = as.numeric(gsub(",",".",pollvalue1)),
    institute = institute
  ) %>%
  select(-pollvalue1)

df <- bind_rows(df1, df2)

df3 <- get_institute_data(institutes[3])

df3 <- df3[c(1,3,4,5,6,7,8,9,13)]

df3 <- df3 %>% 
  gather(key = "party", value = pollvalue, ` CDU/CSU`:` Sonstige`) %>%
  transmute(
    date = as.Date(` `, format = "%d.%m.%Y"),
    party = gsub("\\s","",party),
    pollvalue1 = gsub(" %","",pollvalue),
    pollvalue = as.numeric(gsub(",",".",pollvalue1)),
    institute = institute
  ) %>%
  select(-pollvalue1)

df <- bind_rows(df, df3)

df4 <- get_institute_data(institutes[4])

df4 <- df4[c(1,3,4,5,6,7,8,9,14)]

df4 <- df4 %>% 
  gather(key = "party", value = pollvalue, ` CDU/CSU`:` Sonstige`) %>%
  transmute(
    date = as.Date(` `, format = "%d.%m.%Y"),
    party = gsub("\\s","",party),
    pollvalue1 = gsub(" %","",pollvalue),
    pollvalue = as.numeric(gsub(",",".",pollvalue1)),
    institute = institute
  ) %>%
  select(-pollvalue1)

df <- bind_rows(df, df4)

df5 <- get_institute_data(institutes[5])

df5 <- df5[c(1,3,4,5,6,7,8,9,13)]

df5 <- df5 %>% 
  gather(key = "party", value = pollvalue, ` CDU/CSU`:` Sonstige`) %>%
  transmute(
    date = as.Date(` `, format = "%d.%m.%Y"),
    party = gsub("\\s","",party),
    pollvalue1 = gsub(" %","",pollvalue),
    pollvalue = as.numeric(gsub(",",".",pollvalue1)),
    institute = institute
  ) %>%
  select(-pollvalue1)

df <- bind_rows(df, df5)

df6 <- get_institute_data(institutes[6])

df6 <- df6[c(1,3,4,5,6,7,8,9,10,11,15)]

df6 <- df6 %>% 
  gather(key = "party", value = pollvalue, ` CDU/CSU`:` Sonstige`) %>%
  transmute(
    date = as.Date(` `, format = "%d.%m.%Y"),
    party = gsub("\\s","",party),
    pollvalue1 = gsub(" %","",pollvalue),
    pollvalue = as.numeric(gsub(",",".",pollvalue1)),
    institute = institute
  ) %>%
  select(-pollvalue1)

df <- bind_rows(df, df6)

df7 <- get_institute_data(institutes[7])

df7 <- df7[c(1,3,4,5,6,7,8,9,13)]

df7 <- df7 %>% 
  gather(key = "party", value = pollvalue, ` CDU/CSU`:` Sonstige`) %>%
  transmute(
    date = as.Date(` `, format = "%d.%m.%Y"),
    party = gsub("\\s","",party),
    pollvalue1 = gsub(" %","",pollvalue),
    pollvalue = as.numeric(gsub(",",".",pollvalue1)),
    institute = institute
  ) %>%
  select(-pollvalue1)

df <- bind_rows(df, df7)

rm(df1,df2,df3,df4,df5,df6,df7)
```

```{r}
fnrollmean <- function (x) {
  if (length(x) < 7) {
    rep(NA,length(x)) 
  } else {
    rollmean(x,7,align="center",na.pad=TRUE)
  }
}
```

```{r eval=FALSE, include=FALSE}
drop <- c("FW","PIRATEN","Sonstige")

df.small <- df %>%
  filter(!party %in% drop) %>%
  group_by(date, party) %>%
  mutate(poll_mean = mean(pollvalue, na.rm = TRUE)) %>%
  ungroup() %>%
  arrange(party,date) %>%
  mutate(ma=rollapply(poll_mean,15,mean,align='center',fill=NA))

save(df, df.small, file="polldata.Rda")
```

## Plot
```{r include=FALSE}
library(tidyverse)
library(plotly)
library(viridis)
library(ggthemes)
load("polldata.Rda")
```


```{r fig.height=6, fig.width=10, message=FALSE, warning=FALSE}
df.small %>%
  filter(date > as.Date("2017-01-01")) %>%
  filter(date < as.Date("2018-06-30")) -> df.plot

p <- ggplot(df.plot) +
  geom_point(aes(date, pollvalue, 
                         text = paste("institute:", institute),
                         color = party),
             alpha = 0.6, size = 0.8) +
  geom_line(aes(date, ma, color = party), size = 1) +
  geom_vline(xintercept = as.Date("2017-09-24"), linetype=2) +
  
  geom_vline(xintercept = as.Date("2017-06-01"), linetype=2) +
  geom_vline(xintercept = as.Date("2018-03-01"), linetype=2) +
  scale_color_manual(values = c("#009ee0", "#32302e","#ffed00","#46962b","#ec008c", "#E3000F")) +
  #scale_color_viridis_d(name = NULL) +
  theme_hc() +
  
  labs(x=NULL,y=NULL,color=NULL)

#ggsave(filename = "../figs/polldata.png", width = 7, height = 4)

ggplotly(p)
```

