---
title: "Online media coverage before and after the Bundestag elections in 2017"
author: "Franziska Löw"
date: "06 Sep 2020"
output:
  html_document:
    code_folding: hide
    toc: true
    theme: "lumen"
    highlight: "tango"
    df_print: paged
---

<style type="text/css">
body{ /* Normal  */
      font-size: 18px;
  }
td {  /* Table  */
  font-size: 16px;
}
li {  /* List  */
  font-size: 16px;
}
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r eval=FALSE, include=FALSE}
# options(download.file.method = "wget")
library(RPostgreSQL)
library(tidytext)
library(dplyr)
library(stm)

rm(list = ls())

source("func/functions.R")

color_palette <-c('#d2fbd4','#a5dbc2','#7bbcb0','#559c9e','#3a7c89','#235d72','#123f5a')
```

```{r Connect to PostgresDB, eval=FALSE, include=FALSE}
# create a connection
# loads the PostgreSQL driver
pg <- dbDriver("PostgreSQL")
# creates a connection to the postgres database
pw <- {
  "Up627R&F6"
}

con <- dbConnect(pg,
  user = "Franzi",
  password = pw,
  host = "news-paper.cildx1grqozu.eu-west-2.rds.amazonaws.com",
  port = 5432,
  dbname = "postgres"
)

rm(pw) # removes the password

# check for the news_fulltable
dbExistsTable(con, "news_cleaned")
# TRUE
```

```{r Download from Postgres, eval=FALSE, include=FALSE}
# get News articles
news_df <- dbGetQuery(con, "SELECT * from news_cleaned")
# get press realeases
press_df <- dbGetQuery(con, "SELECT * from press_releases")

# combine both
model_df <- news_df %>%
  dplyr::mutate(
    date = as.Date(date),
    type = "news",
    source = medium
  ) %>%
  select(date, title, title_text, text_length, text_cleaned, type, source) %>%
  bind_rows(press_df %>%
    select(title, date, title_text, type, source, text_length, text_cleaned)) %>%
  dplyr::mutate(
    doc_index = as.numeric(rownames(.)),
    year_week = lubridate::floor_date(date, "1 week")
  )
```

# 1. Introduction

## 1.1 Motivation


**Did media report biased during the election campaign? Have AfD campaigns been increasingly present in the media coverage?**

- It has been argued that the mass media can play a significant role in promoting the success of the populist right. 

  - The mass media are argued to provide legitimacy to the populist right by giving attention to these parties and their issues (Sheets et al. 2016)
  - A higher dose of exposure to populist news coverage enhances both prior agreement and disagreement with populism.
  - Issue ownership: By repeatedly paying attention to specific issues, media raise issue-related public concerns, subsequently increasing citizens' likelihood to cast their vote on the basis of these salient issues (Damstra, 2019).

**Is there a significant difference to the the time period after election?**

- News media are often accused of partisanship and biased reporting, especially during election campaigns (Eberl, 2018)

  - Content analyses in Europe as well as in the United States show that news outlets often differ in their coverage of parties and candidates during election campaigns (Eberl, 2018)

**Theoretical explanation for the disproportionate media coverage of AfD issues.**

- There are structural similarities between the "logic of right-wing populism" and the attention strategy of the mass media. (Gäbler 2017)

  - Populist key messages by political and media actors in news articles provoke more reader comments (Blassnig, 2019). Populist messages often co-occur witha negative, emotionalized, or dramatized communication style (Ernst, Blassnig, Engesser,Büchel, & Esser,2019). Therefore, it can be assumed that the occurrence of populist key messages may have a positive impact on the amount of comments an article receives.
  
  - Populist communication has been described as highly compatible with media logic and attributed with a high news value (Mazzoleni, 2008).
  
  - The media primarily focus on news factors, i.e. the factors that turn an event into news worth reporting. In the case of AfD, a number of news factors are very high: conflict, drama, negativity, surprise, proximity. Finally, the AfD uses the mechanisms of the media attention economy like no other party.
  
  - Media logic (Takens et al. 2013): Media logic refers to 'the news values and the storytelling techniques the media make use of to take advantage of their own medium and its format, and to be competitive in the ongoing struggle to capture people’s attention' (Strömbäck, 2008:233).

## 1.2 Empirical strategy

1. Estimate a **structural topic model** to obtain a topic distribution for each document (press release or news article)

2. Estimate the **cosine similarity** of topic distribution between a document and all documents released +/-4 days.

3. Use topic similarity as the external variable in a regression model: 
  3a) Estimate an **OLS model** for each media outlet $i$

$$
\text{CosineSimilarity}_{i}=\beta_0+\beta_nDn_{j}+\epsilon_i
$$ 
where the $D2_{j}, D3_{j}, ... ,Dn_{j}$ represent dummy variables for a political party $j$

  3b)Estimate a **regression discontinuity model** for each media outlet $i$

$$
\text{CosineSimilarity}_{i}=\beta_0+\beta_1T_i+\beta_nDn_{j}+\beta_{n+n}T_iDn_{j}+\epsilon_i
$$

$$
T_i = 1 \text{ if date } >= \text{election date} \\
T_i = 0 \text{ if date } < \text{election date}
$$
## 1.3 Results

1. **OLS model** (without treatment effect): Considering the whole period, is there a difference between topic similarity of a newspaper and AfD compared to the topic similarity of that newspaper and all other parties? 

-> Yes, looking at the whole time, nearly all newspapers have a significant higher topic similarity with the AfD press releases compared to the other parties.

2. **Regression discont. model** (OLS with treatment effect): Does the topic similarity of a newspaper and the AfD significantly change after the election date?

-> The model shows that for 4 newspapers (Die Welt, stern.de, FOCUS Online, Bild.de), topic similarity with AfD was significantly lower AFTER the election.

The results indicate that - prior to the election - all newspapers (except for Handelsblatt) selected topics more similar to the AfD compared to the other parties. The election significantly changed the topic selection especially for Bild.de as the topic similarity increased significantly for nearly all parties (compared to AfD).

# 3. Calculate STM
```{r Prepare data, eval=FALSE, include=FALSE}
tidy_news_df <- model_df %>%
  select(source, year_week, text_cleaned, doc_index) %>%
  unnest_tokens(word, text_cleaned) %>%
  add_count(word) %>%
  filter(n > 100) %>%
  select(-n)

news_df_sparse <- tidy_news_df %>%
  count(doc_index, word) %>%
  cast_sparse(doc_index, word, n)

## notice that I am scrambling the order of the rows
## the covariate data is in random order now
covariates <- tidy_news_df %>%
  sample_frac() %>%
  distinct(doc_index, source, year_week)
```

```{r Estimate STM, eval=FALSE, include=FALSE}
stmModel <- stm(
  documents = news_df_sparse,
  K = 40, prevalence = ~ source + s(year_week),
  data = covariates, init.type = "Spectral"
)

save(stmModel, covariates, model_df, tidy_news_df, news_df_sparse, file = "model40_weekly_source.Rda")
```

# 3. Calculate cosine similarity

```{r Load and prepare data, include=FALSE}
library(dplyr)
library(stm)
library(tidyr)
library(tidytext)

rm(list = ls())
# 
load("../output/models/model40_weekly_source.Rda")
# 
model_df <- model_df %>%
  dplyr::mutate(
    doc_index = as.numeric(rownames(.)),
    election_dummy = as.factor(ifelse(date <= as.Date("24Sep2017", "%d%b%Y"), "pre", "post")),
    year_week = lubridate::floor_date(date, "1 week")
  )
```

```{r Create matrix of topic distribution, include=FALSE}
td_gamma <- tidy(stmModel,
  matrix = "gamma",
  document_names = rownames(news_df_sparse)
) %>%
  mutate(doc_index = as.integer(document)) %>%
  select(-document)

# convert to "wide" data frame
td_gamma_wide <- td_gamma %>%
  spread(doc_index, gamma) # one row per topic, one column per document

# convert to matrix
gamma <- as.matrix(td_gamma_wide[, -1])
```

```{r Calculate cosine sim}
run_cos_sim_loop <- function(target) {
  small_df <- model_df %>% filter(source == target)
  doc_indices <- small_df$doc_index

  print(paste("Number of documents:", length(doc_indices)))

  rm(cosine_distances)
  i <- 1
  for (x in doc_indices) {
    print(i)
    temp <- calc_cos_sim(x, gamma)

    # append the results to a data frame
    cosine_distances <- if (!exists("cosine_distances")) temp else rbind(cosine_distances, temp)

    i <- i + 1
  }

  cosine_distances
}
```

```{r eval=FALSE, include=FALSE}
# DIE WELT
target_source <- "DIE WELT"
welt <- run_cos_sim_loop(target_source)
save(welt, file = paste0("../output/cosine_dist_", target_source, ".Rda"))

# SPIEGEL ONLINE
target_source <- "SPIEGEL ONLINE"
spiegel <- run_cos_sim_loop(target_source)
save(spiegel, file = paste0("../output/cosine_dist_", target_source, ".Rda"))

# stern.de
target_source <- "stern.de"
stern <- run_cos_sim_loop(target_source)
save(stern, file = paste0("../output/cosine_dist_", target_source, ".Rda"))

# ZEIT ONLINE
target_source <- "ZEIT ONLINE"
zeit <- run_cos_sim_loop(target_source)
save(zeit, file = paste0("../output/cosine_dist_", target_source, ".Rda"))

# Handelsblatt
target_source <- "Handelsblatt"
handelsblatt <- run_cos_sim_loop(target_source)
save(handelsblatt, file = paste0("../output/cosine_dist_", target_source, ".Rda"))

# FOCUS Online
target_source <- "FOCUS Online"
focus <- run_cos_sim_loop(target_source)
save(focus, file = paste0("../output/cosine_dist_", target_source, ".Rda"))

# Bild.de
target_source <- "Bild.de"
bild <- run_cos_sim_loop(target_source)
save(bild, file = paste0("../output/cosine_dist_", target_source, ".Rda"))
```

# 4. Estimate OLS

## 4.2. Regression discontinuity model

The idea of regression regression discontinuity design is to use observations with a $W_i$ close to $c$ for the estimation of $\beta_1$. $\beta_1$ is the average treatment effect for observations with $W_i = c$ which is assumed to be a good approximation to the overall treatment effect. In other words, $\beta_1$ gives us the average change of news coverage of media outlet $i$ after the election day. Since interaction terms $T_iDn_{j}$ are included, we can estimate the treatment effect for each party. 

Calculate a regression discontinuity model for each newspaper $i$.

$$
\text{CosineSimilarity}_{i}=\beta_0+\beta_1T_i+\beta_2W_{centered}+\beta_nDn_{j}+\beta_{n+n}T_iDn_{j}+\epsilon_i
$$

where $D2_{j}, D3_{j}, ... ,Dn_{j}$ represent dummy variables for a political party $j$ and $W_{centered} = $ date - election day. 

$$
T_i = 1 \text{ if date } >= \text{election date} \\
T_i = 0 \text{ if date } < \text{election date}
$$

so that the receipt of treatment $T_i$ is determined by the threshold $c$ (election day) of the continuous variable $W_i$ (date), the so called running variable. 

```{r Plot datapoints}
plot_data <- function(dataframe) {
  cosine_distances_df %>%
    filter(type2 == "press") %>%
    ggplot(aes(date1, cos_sim, color = election_dummy)) +
    geom_point(show.legend = F) +
    geom_vline(xintercept = election_date) +
    ggthemes::theme_clean() +
    labs(y = "Cosine similarity of topic distribution", x = "") +
    facet_wrap(~source2)
}
```

```{r Calculate models}
calc_rd_dummy<- function(dataframe, target_source) { 
  temp_df <- dataframe %>%
    filter(type2 == "press") %>%
    mutate(
      X_centered = I(date1 - election_date),
      treatedTRUE = ifelse(date1 >= election_date, 1, 0))

  model_outcome <- temp_df %$%
    lm(log(cos_sim) ~ treatedTRUE + source2 + X_centered)
}

calc_rd_dummy_interaction <- function(dataframe, target_source) { 
  temp_df <- dataframe %>%
    filter(type2 == "press") %>%
    mutate(
      X_centered = I(date1 - election_date),
      treatedTRUE = ifelse(date1 >= election_date, 1, 0))

  model_outcome <- temp_df %$%
    lm(log(cos_sim) ~ treatedTRUE * source2 + X_centered)
}
```

```{r message=FALSE}
target_source <- "DIE WELT"
#load(paste0("../output/cosine_dist_", target_source, ".Rda"))
cosine_distances_df <- df_prep(welt)
welt_model1 <- calc_rd_dummy(cosine_distances_df, target_source)
welt_model2 <- calc_rd_dummy_interaction(cosine_distances_df, target_source)

target_source <- "stern.de"
#load(paste0("../output/cosine_dist_", target_source, ".Rda"))
cosine_distances_df <- df_prep(stern)
stern_model1 <- calc_rd_dummy(cosine_distances_df, target_source)
stern_model2 <- calc_rd_dummy_interaction(cosine_distances_df, target_source)

target_source <- "ZEIT ONLINE"
#load(paste0("../output/cosine_dist_", target_source, ".Rda"))
cosine_distances_df <- df_prep(zeit)
zeit_model1 <- calc_rd_dummy(cosine_distances_df, target_source)
zeit_model2 <- calc_rd_dummy_interaction(cosine_distances_df, target_source)

target_source <- "Handelsblatt"
#load(paste0("../output/cosine_dist_", target_source, ".Rda"))
cosine_distances_df <- df_prep(handelsblatt)
handelsblatt_model1 <- calc_rd_dummy(cosine_distances_df, target_source)
handelsblatt_model2 <- calc_rd_dummy_interaction(cosine_distances_df, target_source)

target_source <- "FOCUS Online"
#load(paste0("../output/cosine_dist_", target_source, ".Rda"))
cosine_distances_df <- df_prep(focus)
focus_model1 <- calc_rd_dummy(cosine_distances_df, target_source)
focus_model2 <- calc_rd_dummy_interaction(cosine_distances_df, target_source)

target_source <- "Bild.de"
#load(paste0("../output/cosine_dist_", target_source, ".Rda"))
cosine_distances_df <- df_prep(bild)
bild_model1 <- calc_rd_dummy(cosine_distances_df, target_source)
bild_model2 <- calc_rd_dummy_interaction(cosine_distances_df, target_source)

target_source <- "SPIEGEL ONLINE"
#load(paste0("../output/cosine_dist_", target_source, ".Rda"))
cosine_distances_df <- df_prep(spiegel)
spiegel_model1 <- calc_rd_dummy(cosine_distances_df, target_source)
spiegel_model2 <- calc_rd_dummy_interaction(cosine_distances_df, target_source)
```

### 4.1.1. Without interaction terms

$$
\text{CosineSimilarity}_{i}=\beta_0+\beta_1T_i+\beta_nDn_{j}+\epsilon_i
$$

```{r Model reults without interaction terms, message=FALSE, warning=FALSE}
stargazer::stargazer(welt_model1, stern_model1, zeit_model1, handelsblatt_model1, focus_model1, bild_model1, spiegel_model1,
                     column.labels = c("DIE WELT", "stern.de", "ZEIT ONLINE", "Handelsblatt", "FOCUS Online", "Bild.de", "SPIEGEL ONLINE"),
                     dep.var.labels = paste0("Cosine similarity of topic distribution"), 
                     single.row = TRUE,
                     #font.size = "tiny",
                     type = "text")
```

#### Interpretation:

**DIE WELT**

- Treatment effect: After the election, cosine similarity of news coverage drops by ~10,8% on average. 
- Dummy variables: Compared to AfD, news coverage with all other parties is less similar

**stern.de**

- Treatment effect: After the election, cosine similarity of news coverage drops by 16,2% on average. 
- Dummies: Compared to AfD, news coverage with all other parties is less similar

**Handelsblatt**

- Treatment effect: After the election, cosine similarity of news coverage drops by 14,6% on average. 
- Dummies: No significant difference


### 4.1.2. With interaction terms

$$
\text{CosineSimilarity}_{i}=\beta_0+\beta_1T_i+\beta_2Dn_{j}+\beta_3T_iDn_{j}+\epsilon_i
$$

The interaction term $T_i*Dn$ means that the slope can vary on either side of the treatment threshold for each party. 

- The coefficient $\beta_1$ is how the intercept jumps (the RDD effect)
- $\beta_3$ is how the slope changes for each party

```{r Model reults with interaction terms, message=FALSE, warning=FALSE}
stargazer::stargazer(welt_model2, stern_model2, zeit_model2, handelsblatt_model2, focus_model2, bild_model2, spiegel_model2,
                     column.labels = c("DIE WELT", "stern.de", "ZEIT ONLINE", "Handelsblatt", "FOCUS Online", "Bild.de", "SPIEGEL ONLINE"),
                     dep.var.labels = paste0("Cosine similarity of topic distribution"), 
                     single.row = TRUE,
                     #font.size = "tiny",
                     type = "text")
```

#### Interpretation: 

**DIE WELT**

- Treatment effect: Topic similarity with AfD decreased by 14,7%

**stern.de**

- Treatment effect: Topic similarity with AfD decreased by 16,9% 

**ZEIT ONLINE**

**Handelsblatt**

**FOCUS Online**

- Treatment effect: Topic similarity with AfD decreased by 11,1%

- Dummy variables: The cosine similarity is on average less for all parties compared to AfD *BEFORE* the election day. E.g. compared to the AfD, the cosine similarity of topics is ~33% lower for B90/Grüne, meaning that topic coverage of Bild.de is 33% lower for B90/Grüne compared to AfD before the election day.

- Interaction terms

$$
ln(\text{CosineSimilarity}_{FOCUS}(D_{B90}=1))=\beta_0=\beta_1(-0.118)*T+\beta_2+\beta_3(0.138)*T \\
ln(\text{CosineSimilarity}_{FOCUS}(D_{B90}=1))= (\beta_0-\beta_2)+(-0.118+0.138)*T
$$

$$
ln(\text{CosineSimilarity}_{FOCUS}(D_{B90}=0))=\beta_0-0.118*T
$$
When $D_{B90}$ switches from 0 to 1, the coefficient of the treatment effect $T$ increases by 0.138.

For $D_{B90}=0$ (treatment effect on AfD): $100(exp(-0.118)-1) = -11,1%$
--> The election day had a negative impact on the topic similarity of FOCUS and AfD of ~11%

For $D_{B90}=1$ (treatment effect on B90/Grüne): $100(exp(-0.118+0.138)-1) = 2%$
--> The election day had a positive impact on the topic similarity of FOCUS and B90/Grüne (compared to AfD) of ~2%

#### Compare models

**DIE WELT**
```{r anova DIE WLET}
anova(welt_model1, welt_model2)
```

The result shows a Df of 5 (indicating that the more complex model has 5 additional parameter) and a p-value of 0.42. This means that adding the interaction terms to the model did NOT lead to a significantly improved fit over the model.(Do the same check for all other media outlet models)